Model Checking Differentially Private Properties

### 2018-02-01 dpCTL.tex
minor fixes

The logic $\dpCTLstar$ is designed to specify differentially private
mechanisms. We introduce the differentially private path quantifer
$\dpriv{\epsilon}{\delta}$ in $\dpCTLstar$. For any path formula
$\phi$, a state $s$ in a Markov chain $K$ satisfies
$\dpriv{\epsilon}{\delta} \phi$ if the probability of having $\phi$-paths
from $s$ is close to the probabilities of
having $\phi$-paths from its neighboring states.
%We give semantics for both Markov chains and Markov decision processes.

>>>>

The logic $\dpCTLstar$ is designed to specify differentially private
mechanisms. We introduce the differentially private path quantifer
$\dpriv{\epsilon}{\delta}$ in $\dpCTLstar$. For any path formula
$\phi$, a state $s$ in a Markov chain $K$ satisfies
$\dpriv{\epsilon}{\delta} \phi$ if the probability of having $\phi$-paths
from $s$ is close to the probabilities of
having $\phi$-paths from its neighbors.
%We give semantics for both Markov chains and Markov decision processes.

### 2018-02-01 examples.tex
minor fixes

\subsection{Survey Mechanisms}
\label{subsec:survey}

Consider the survey question: have you been diagnosed
with the disease $X$? In order to protect surveyees' privacy, each
answers the question as follows. The surveyee first flips a
coin. If it is tail, she answers the question truthfully. Otherwise,
she randomly answers \textit{Yes} or \textit{No}
uniformly~\cite{DR:14:AFDP}.

Let us analyze the mechanism briefly. The data universe $\mathcal{X}$
is $\{ +, - \}$. The mechanism $M$ is a randomized algorithm with
inputs in $\mathcal{X}$ and outputs in $\{ Y, N \}$. For any $x \in
\mathcal{X}$, we have $\frac{1}{4} \leq \Pr[M (x) = Y] \leq
\frac{3}{4}$. Hence $\Pr[M (x) = Y] \leq \frac{3}{4} = 3 \cdot
\frac{1}{4} \leq e^{\ln 3} \Pr[M (x') = Y]$ for any neighboring $x, x'
\in \mathcal{X}$. Similarly, $\Pr[M (x) = N] \leq e^{\ln 3} \Pr[M (x')
= N]$. The survey mechanism is $(\ln 3, 0)$-differentially private.

@@

\begin{example}\label{exa:survey}
  Figure~\ref{figure:2-dp} shows the survey
mechanism and its corresponding Markov chain.
In the figure, the states
$+$ and $-$ denote the diagnoses are positive or negative
respectively; the states $s$ and $t$ denote answers to the survey
question and hence let $\mathit{out}_Y \in L (s)$ and $\mathit{out}_N
\in L (t)$.
States $+$ and $-$ are neighbors.
The random noise boosts the probability of answering
\textit{Yes} or \textit{No} to at least $1/4$ regardless of
diagnoses. Inferences on individual diagnosis can be plausibly denied.
Missing transitions (such as those from $s$ and $t$) lead
to a special state $\dagger$ with a self-loop. We omit the transitions
and the state $\dagger$ for clarity.
\end{example}

\subsection{Truncated $\alpha$-Geometric Mechanism}\label{subsec:geometric}
More sophisticated differentially private mechanisms are
available. Consider a query
$f : \mathcal{X}^n \rightarrow \bbfZ^{\geq 0}$. Let $\alpha \in (0, 1)$.
The \emph{$\alpha$-geometric mechanism}
outputs $f(x) + Y$ where $Y$ is a random variable with the geometric
distribution~\cite{GRS:09:UUPM,GRS:12:UUPM} :
\[
\Pr[Y = y] = \frac{1 - \alpha}{1 + \alpha}\alpha^{|y|}
\textmd{ for } y \in \bbfZ.
\]
The $\alpha$-geometric mechanism is oblivious since it has the same
output distribution on any inputs $x, x'$ with $f (x) = f
(x')$. Moreover, it is $(- {\Delta (f)} \ln \alpha, 0)$-differentially
private for any query $f$ with sensitivity $\Delta (f)$. The range of
the mechanism
however is $\bbfZ$. It may give nonsensical outputs such as
negative integers for queries with positive ranges.

@@

\begin{figure}[tbh]
  \centering
  \begin{subfigure}{.40\columnwidth}
    \[
    \begin{array}{|c|c|c|c|c|c|c|}
      \hline
      \textmd{input/output} & 0 & 1 & 2 & 3 & 4 & 5\\
      \hline
      0 & {2}/{3}  & {1}/{6} & {1}/{12} & {1}/{24}  & {1}/{48} & {1}/{48}\\
      \hline
      1 & {1}/{3}  & {1}/{3} & {1}/{6} & {1}/{12}  & {1}/{24} & {1}/{24}\\
      \hline
      2 & {1}/{6}  & {1}/{6} & {1}/{3} & {1}/{6}   & {1}/{12} & {1}/{12}\\
      \hline
      3 & {1}/{12} & {1}/{12} & {1}/{6} & {1}/{3}   & {1}/{6} & {1}/{6}\\
      \hline
      4 & {1}/{24} & {1}/{24} & {1}/{12} & {1}/{6}   & {1}/{3} & {1}/{3}\\
      \hline
      5 & {1}/{48} & {1}/{48} & {1}/{24} & {1}/{12} & {1}/{6} & {2}/{3}\\
      \hline
    \end{array}
    \]
    \caption{$1/2$-Geometric Mechanism}
    \label{figure:geometric-mechanism-table}
  \end{subfigure}
  \hspace{.08\columnwidth}
  \begin{subfigure}{.50\columnwidth}
    \centering
  \resizebox{.85\columnwidth}{!}{
    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
      distance=2cm,node/.style={circle,draw}]
      \node[node] (i0) at (0,  1) { $s_0$ };
      \node at (0, 1.5) { $\mathit{in}_0$ };
      \node at (0, 0) { $\vdots$ };
      \node at (.8, -.8) { $\vdots$ };
      \node[node] (i5) at (0, -1) { $s_5$ };
      \node at (0, -.5) { $\mathit{in}_5$ };

@@

Similar to the survey mechanism, it is
straightforward to model the truncated $\frac{1}{2}$-geometric
mechanisms as Markov chains.

\begin{example}
Let the state $s_k$ and $t_l$
denote the input value $k$ and output value $l$ respectively. Define
the state set $S = \{ s_k, t_k : k \in \{ 0, 1, \ldots, n \} \}$.
The probability transition $\wp (s_k, t_l)$ is the probability of the
output value $l$ on the input value $k$ as defined in the
mechanism. Moreover, we have $\mathit{in}_k \in L (s_k)$ and
$\mathit{out}_k \in L (t_k)$ for $k \in \{ 0, 1, \ldots, n \}$.
If $\Delta (f) = 1$, $s_k$ and $s_l$ are
neighbors iff $| k - l | \leq 1$.
Figure~\ref{figure:geometric-mechanism-markov-chain} gives
the Markov chain of the truncated
$\frac{1}{2}$-geometric mechanism over $\{ 0, 1, \ldots, 5 \}$.
\end{example}

>>>>

\subsection{Survey Mechanism}
\label{subsec:survey}

Consider the survey question: have you been diagnosed
with the disease $X$? In order to protect surveyees' privacy, each
answers the question as follows. The surveyee first flips a
coin. If it is tail, she answers the question truthfully. Otherwise,
she randomly answers \textit{Yes} or \textit{No}
uniformly (Figure~\ref{figure:2-dp-table})~\cite{DR:14:AFDP}.

Let us analyze the mechanism briefly. The data universe $\mathcal{X}$
is $\{ +, - \}$. The mechanism $M$ is a randomized algorithm with
inputs in $\mathcal{X}$ and outputs in $\{ Y, N \}$. For any $x \in
\mathcal{X}$, we have $\frac{1}{4} \leq \Pr[M (x) = Y] \leq
\frac{3}{4}$. Hence $\Pr[M (x) = Y] \leq \frac{3}{4} = 3 \cdot
\frac{1}{4} \leq e^{\ln 3} \Pr[M (x') = Y]$ for any neighboring $x, x'
\in \mathcal{X}$. Similarly, $\Pr[M (x) = N] \leq e^{\ln 3} \Pr[M (x')
= N]$. The survey mechanism is $(\ln 3, 0)$-differentially private.
The random noise boosts the probability of answering
\textit{Yes} or \textit{No} to at least $1/4$ regardless of
diagnoses. Inferences on individual diagnosis can be plausibly denied.

@@

Figure~\ref{figure:2-dp-mdp} shows the corresponding Markov chain.
In the figure, the states
$+$ and $-$ denote the diagnoses are positive or negative
respectively; the states $s$ and $t$ denote answers to the survey
question and hence let $\mathit{out}_Y \in L (s)$ and $\mathit{out}_N
\in L (t)$.
States $+$ and $-$ are neighbors.
Missing transitions (such as those from $s$ and $t$) lead
to a special state $\dagger$ with a self-loop. We omit the transitions
and the state $\dagger$ for clarity.

\subsection{Truncated $\alpha$-Geometric Mechanism}\label{subsec:geometric}
More sophisticated differentially private mechanisms are
available. Consider a query
$f : \mathcal{X}^n \rightarrow \bbfZ^{\geq 0}$. Let $\alpha \in (0, 1)$.
The \emph{$\alpha$-geometric mechanism}
outputs $f(x) + Y$ where $Y$ is a random variable with the geometric
distribution~\cite{GRS:09:UUPM,GRS:12:UUPM} :
\[
\Pr[Y = y] = \frac{1 - \alpha}{1 + \alpha}\alpha^{|y|}
\textmd{ for } y \in \bbfZ.
\]
The $\alpha$-geometric mechanism is oblivious since it has the same
output distribution on any inputs $x, x'$ with $f (x) = f
(x')$. It is $(- {\Delta (f)} \ln \alpha, 0)$-differentially
private for any query $f$ with sensitivity $\Delta (f)$. The range of
the mechanism
however is $\bbfZ$. It may give nonsensical outputs such as
negative integers for queries with positive ranges.

@@

\begin{figure}[tbh]
  \centering
  \begin{subfigure}{.40\columnwidth}
    \[
    \begin{array}{|c|c|c|c|c|c|c|}
      \hline
      \textmd{input/output} & 0 & 1 & 2 & 3 & 4 & 5\\
      \hline
      0 & {2}/{3}  & {1}/{6} & {1}/{12} & {1}/{24}  & {1}/{48} & {1}/{48}\\
      \hline
      1 & {1}/{3}  & {1}/{3} & {1}/{6} & {1}/{12}  & {1}/{24} & {1}/{24}\\
      \hline
      2 & {1}/{6}  & {1}/{6} & {1}/{3} & {1}/{6}   & {1}/{12} & {1}/{12}\\
      \hline
      3 & {1}/{12} & {1}/{12} & {1}/{6} & {1}/{3}   & {1}/{6} & {1}/{6}\\
      \hline
      4 & {1}/{24} & {1}/{24} & {1}/{12} & {1}/{6}   & {1}/{3} & {1}/{3}\\
      \hline
      5 & {1}/{48} & {1}/{48} & {1}/{24} & {1}/{12} & {1}/{6} & {2}/{3}\\
      \hline
    \end{array}
    \]
    \caption{$\frac{1}{2}$-Geometric Mechanism}
    \label{figure:geometric-mechanism-table}
  \end{subfigure}
  \hspace{.08\columnwidth}
  \begin{subfigure}{.50\columnwidth}
    \centering
  \resizebox{.85\columnwidth}{!}{
    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
      distance=2cm,node/.style={circle,draw}]
      \node[node] (i0) at (0,  1) { $s_0$ };
      \node at (0, 1.5) { $\mathit{in}_0$ };
      \node at (0, 0) { $\vdots$ };
      \node at (.8, -.8) { $\vdots$ };
      \node[node] (i5) at (0, -1) { $s_5$ };
      \node at (0, -.5) { $\mathit{in}_5$ };

@@

Similar to the survey mechanism, it is
straightforward to model the truncated $\frac{1}{2}$-geometric
mechanisms as Markov chains. Let the state $s_k$ and $t_l$
denote the input value $k$ and output value $l$ respectively. Define
the state set $S = \{ s_k, t_k : k \in \{ 0, 1, \ldots, n \} \}$.
The probability transition $\wp (s_k, t_l)$ is the probability of the
output value $l$ on the input value $k$ as defined in the
mechanism. Moreover, we have $\mathit{in}_k \in L (s_k)$ and
$\mathit{out}_k \in L (t_k)$ for $k \in \{ 0, 1, \ldots, n \}$.
If $\Delta (f) = 1$, $s_k$ and $s_l$ are
neighbors iff $| k - l | \leq 1$.
Figure~\ref{figure:geometric-mechanism-markov-chain} gives
the Markov chain of the truncated
$\frac{1}{2}$-geometric mechanism over $\{ 0, 1, \ldots, 5 \}$.

### 2018-02-01 introduction.tex
minor fixes

\noindent
\emph{Related Work.}
An early attempt on formal verification of differential privacy
is~\cite{TKD:11:FVDPIS}. The work formalizes differential privacy in
the framework of information leakage. The connection between
differential privacy and information leakage is investigated
in~\cite{AACDP:15:OILDPM,GMP:16:PDPFPS}. Type systems for differential
privacy have been developed
in~\cite{WHRP:17:FADP,GHHNP:13:LDPDP,RP:10:DMTGS}. A light-weight
technique for checking differential privacy can be found
in~\cite{ZK:17:LTADPP}. Lots of formal Coq proofs in differential
privacy are reported
in~\cite{BFGGHS:16:APCDP,BGGHS:16:PDPPC,BKOB:12:PRRDP,BDGKZ:13:VCDPASM,BGAHKS:14:PDPHL,BGAHRS:15:HOARRT,BFGAGHS:16:DPBP}. This
work emphasizes on model checking differential privacy. We develop a
framework to formalize and analyze differential privacy in Markov
chains and Markov decision processes.

\noindent
\emph{Contributions.} Our main contributions are threefold. (I)
We introduce the logic $\dpCTLstar$ for reasoning about properties in
differential privacy. We use the logic to express classical
differential privacy properties, and moreover, complex properties such as those
involving infinite behaviors can also be expressed. (II) We model several differentially private mechanisms, including stream processing, in Markov chains or Markov decision processes.
(III) We show that the model checking
problem for Markov chains is standard. For Markov decision processes we show that it is undecidable.

>>>>

\noindent
\emph{Related Work.}
An early attempt on formal verification of differential privacy
is~\cite{TKD:11:FVDPIS}. The work formalizes differential privacy in
the framework of information leakage. The connection between
differential privacy and information leakage is investigated
in~\cite{AACDP:15:OILDPM,GMP:16:PDPFPS}. Type systems for differential
privacy have been developed
in~\cite{WHRP:17:FADP,GHHNP:13:LDPDP,RP:10:DMTGS}. A light-weight
technique for checking differential privacy can be found
in~\cite{ZK:17:LTADPP}. Lots of formal Coq proofs about differential
privacy are reported
in~\cite{BFGGHS:16:APCDP,BGGHS:16:PDPPC,BKOB:12:PRRDP,BDGKZ:13:VCDPASM,BGAHKS:14:PDPHL,BGAHRS:15:HOARRT,BFGAGHS:16:DPBP}. This
work emphasizes on model checking differential privacy. We develop a
framework to formalize and analyze differential privacy in Markov
chains and Markov decision processes.

\noindent
\emph{Contributions.} Our main contributions are threefold. (I)
We introduce the logic $\dpCTLstar$ for reasoning about properties in
differential privacy. We use the logic to express classical
differential privacy properties, and moreover, complex properties such as those
involving infinite behaviors can also be expressed. (II) We model several differentially private mechanisms, including stream processing, in Markov chains or Markov decision processes.
(III) We show that the model checking
problem for Markov chains is standard. For Markov decision processes, we show that it is undecidable.

### 2018-02-01 semantics.tex
bib
Correct some spelling errors and others
minor fixes

Moreover, the relation $K, \neighbor{S}, \pi \models \phi$ is defined as
for the standard LTL formulas~\cite{which-to-cite?}. We only recall the semantics for $\X$ and $\until$ as follows:
\begin{eqnarray*}
  K, \neighbor{S}, \pi \models \X \phi
  & \textmd{ if } &
  K, \neighbor{S}, \pi[1] \models \phi\\
  K, \neighbor{S}, \pi \models \phi \until \psi
  & \textmd{ if } &
  \textmd{there is a } j \geq 0 \textmd{ such that }
  K, \neighbor{S}, \pi[j] \models \psi \textmd{ and } \\
  & & K, \neighbor{S}, \pi[k] \models \phi
      \textmd{ for every } 0 \leq k < j
\hide{
  K, \neighbor{S}, \pi \models \Phi \buntil{n} \Psi
  & \textmd{ if } &
  \textmd{there is a } 0 \leq j \leq n \textmd{ such that }
  K, \neighbor{S}, \pi[j] \models \Psi \textmd{ and } \\
  & & K, \neighbor{S}, \pi[k] \models \Phi
      \textmd{ for every } 0 \leq k < j
}
\end{eqnarray*}

@@

Other than the differentially private operator, the semantics of
$\dpCTLstar$ is standard~\cite{BK:08:PMC}.
To intuit the semantics of $\dpriv{\epsilon}{\delta} \phi$,
recall that  $\myPr{s}{K}{N}{\phi}$ is the probability of having
paths satisfying $\phi$ from $s$. Thus, a state $s$ satisfies
$\dpriv{\epsilon}{\delta} \phi$ if having paths satisfying $\phi$
from $s$ is $(\epsilon, \delta)$-close to having paths satisfying
$\phi$ from every neighbor of $s$. In other words, it is
probabilistically similar to observe paths satisfying $\phi$ from $s$
and from all neighbors of $s$.

\subsection{Model Checking Algorithms}
We first describe the model checking algorithm for $\dpCTL$, which follows the classical
algorithms for $\PCTL$ by computing the states satisfying
sub state-formulae inductively~\cite{HanssonJ94,BK:08:PMC}. It hence suffices to
consider the inductive step where the states satisfying the subformula
$\dpriv{\epsilon}{\delta} (\phi)$ is computed.

@@

In the classical $\PCTL$ model checking algorithm for Markov chains,
states satisfying the subformula $\PJ{J} \phi$ are obtained by
computing $\myPr{s}{K}{\neighbor{S}}{\phi}$ for state $s \in S$.
These probabilities can be computed by solving linear equations or
through iterative approximations. We summarize it in the following
theorem (details see~\cite{BK:08:PMC}):

@@

In order to determine the set $\{ s : K, \neighbor{S}, s \models
\dpriv{\epsilon}{\delta} \varphi \}$, our algorithm first computes
the probabilities $p (s) = \myPr{s}{K}{\neighbor{S}}{\varphi}$ for every
$s \in S$ (Algorithm~\ref{algorithm:sat-dpriv-dtmc}). For each $s \in S$,
it then compares the probabilities $p (s)$ and $p (t)$ for every
neighbor $t$ of $s$. If there is a neighbor $t$ such that $p (s)$ and
$p (t)$ are not $(\epsilon, \delta)$-close, the state $s$
is removed from the result. Hence
Algorithm~\ref{algorithm:sat-dpriv-dtmc} returns all states which are
$(\epsilon, \delta)$-close to their neighbors.
The algorithm requires at most $O
(|S|^2)$ additional steps. 
We hence have the following results:

\begin{algorithm}[tbh]
  \begin{algorithmic}[1]
    \Function{SAT}{$K$, $\neighbor{S}$, $\phi$}
    \Match{$\phi$}
    \Comment{by Lemma~\ref{lemma:PJ-subroutines}}
    \Case{$\X \Psi$}
      \State{$B \leftarrow \textmd{SAT} (K, \neighbor{S}, \Psi)$}
      \State{$p(s) \leftarrow \myPr{s}{K}{\neighbor{S}}{\gX B}$
        for every $s \in S$}
    \EndCase
    \Case{$\Psi \until \Psi'$}
      \State{$B \leftarrow \textmd{SAT} (K, \neighbor{S}, \Psi)$}
      \State{$C \leftarrow \textmd{SAT} (K, \neighbor{S}, \Psi')$}
      \State{$p(s) \leftarrow \myPr{s}{K}{\neighbor{S}}{B \guntil C}$ for
        every $s \in S$}
    \EndCase
    \EndMatch
    \State{$R \leftarrow S$}
    \For{$s \in S$}
      \For{$t$ with $s \neighbor{K} t$}
        \If{$p(s) \not\leq e^{\epsilon} p(t) + \delta$ or
            $p(t) \not\leq e^{\epsilon} p(s) + \delta$}
        {remove $s$ from $R$}
        \EndIf
      \EndFor
    \EndFor

    \State{\Return $R$}
    \EndFunction
  \end{algorithmic}
  \caption{SAT($\phi$, $\neighbor{K}$)}
  \label{algorithm:sat-dpriv-dtmc}
\end{algorithm}

>>>>

Moreover, the relation $K, \neighbor{S}, \pi \models \phi$ is defined as
for the standard LTL formulas~\cite{ManneP92}. We only recall the semantics for $\X$ and $\until$ as follows:
\begin{eqnarray*}
  K, \neighbor{S}, \pi \models \X \phi
  & \textmd{ if } &
  K, \neighbor{S}, \pi[1] \models \phi\\
  K, \neighbor{S}, \pi \models \phi \until \psi
  & \textmd{ if } &
  \textmd{there is a } j \geq 0 \textmd{ such that }
  K, \neighbor{S}, \pi[j] \models \psi \textmd{ and } \\
  & & K, \neighbor{S}, \pi[k] \models \phi
      \textmd{ for every } 0 \leq k < j
\hide{
  K, \neighbor{S}, \pi \models \Phi \buntil{n} \Psi
  & \textmd{ if } &
  \textmd{there is a } 0 \leq j \leq n \textmd{ such that }
  K, \neighbor{S}, \pi[j] \models \Psi \textmd{ and } \\
  & & K, \neighbor{S}, \pi[k] \models \Phi
      \textmd{ for every } 0 \leq k < j
}
\end{eqnarray*}

@@

Other than the differentially private operator, the semantics of
$\dpCTLstar$ is standard~\cite{BK:08:PMC}.
To intuit the semantics of $\dpriv{\epsilon}{\delta} \phi$,
recall that  $\myPr{s}{K}{N}{\phi}$ is the probability of having
paths satisfying $\phi$ from $s$. A state $s$ satisfies
$\dpriv{\epsilon}{\delta} \phi$ if having paths satisfying $\phi$
from $s$ is $(\epsilon, \delta)$-close to having paths satisfying
$\phi$ from every neighbor of $s$. Informally, it is
probabilistically similar to observe paths satisfying $\phi$ from $s$
and from all its neighbors.

\subsection{Model Checking Algorithms}
We describe the model checking algorithm for $\dpCTL$. The algorithm follows the classical
algorithms for $\PCTL$ by computing the states satisfying
sub state-formulae inductively~\cite{HanssonJ94,BK:08:PMC}. It hence suffices to
consider the inductive step where the states satisfying the subformula
$\dpriv{\epsilon}{\delta} (\phi)$ are to be computed.

@@

In the classical $\PCTL$ model checking algorithm for Markov chains,
states satisfying the subformula $\PJ{J} \phi$ are obtained by
computing $\myPr{s}{K}{\neighbor{S}}{\phi}$ for state $s \in S$.
These probabilities can be obtained by solving linear equations or
through iterative approximations. We summarize it in the following
theorem (details see~\cite{BK:08:PMC}):

@@

In order to determine the set $\{ s : K, \neighbor{S}, s \models
\dpriv{\epsilon}{\delta} \varphi \}$, our algorithm computes
the probabilities $p (s) = \myPr{s}{K}{\neighbor{S}}{\varphi}$ for every
$s \in S$ (Algorithm~\ref{algorithm:sat-dpriv-dtmc}). For each $s \in S$,
it then compares the probabilities $p (s)$ and $p (t)$ for every
neighbor $t$ of $s$. If there is a neighbor $t$ such that $p (s)$ and
$p (t)$ are not $(\epsilon, \delta)$-close, the state $s$
is removed from the result. 
Algorithm~\ref{algorithm:sat-dpriv-dtmc} returns all states which are
$(\epsilon, \delta)$-close to their neighbors.
The algorithm requires at most $O
(|S|^2)$ additional steps.
We hence have the following results:

\begin{algorithm}[tbh]
  \begin{algorithmic}[1]
    \Procedure{SAT}{$K$, $\neighbor{S}$, $\phi$}
    \Match{$\phi$}
    \Comment{by Lemma~\ref{lemma:PJ-subroutines}}
    \Case{$\X \Psi$}
      \State{$B \leftarrow \textmd{SAT} (K, \neighbor{S}, \Psi)$}
      \State{$p(s) \leftarrow \myPr{s}{K}{\neighbor{S}}{\gX B}$
        for every $s \in S$}
    \EndCase
    \Case{$\Psi \until \Psi'$}
      \State{$B \leftarrow \textmd{SAT} (K, \neighbor{S}, \Psi)$}
      \State{$C \leftarrow \textmd{SAT} (K, \neighbor{S}, \Psi')$}
      \State{$p(s) \leftarrow \myPr{s}{K}{\neighbor{S}}{B \guntil C}$ for
        every $s \in S$}
    \EndCase
    \EndMatch
    \State{$R \leftarrow S$}
    \For{$s \in S$}
      \For{$t$ with $s \neighbor{S} t$}
        \If{$p(s) \not\leq e^{\epsilon} p(t) + \delta$ or
            $p(t) \not\leq e^{\epsilon} p(s) + \delta$}
        {remove $s$ from $R$}
        \EndIf
      \EndFor
    \EndFor

    \State{\Return $R$}
    \EndProcedure
  \end{algorithmic}
  \caption{SAT($K$, $\neighbor{S}$, $\phi$)}
  \label{algorithm:sat-dpriv-dtmc}
\end{algorithm}
